---
output:
  pdf_document: default
  html_document: default
geometry: margin=2cm
---
<!-- This is an RMarkdown document that reproduces the statistics and graphs
from Tilt Processing.R. It loads up an archive dataset generated from
the Tilt Database Cleanup.R script -->



```{r include = FALSE}
# Setup variables for the rest of the script.

# First we set the place we want the output PDF to be saved.
# Set the working directory for saving files.
setwd("~/Google Drive/Beer Things/")

# Which data file we want to use for the flyer. The assumption is that this data
# came from Tilt Database Cleanup.R and is ready to be used.
# Load and transform archival data.
read.csv(file = "~/Google Drive/Beer Things/Tilt Archive/2018-04_Munich Helles",
         stringsAsFactors = FALSE) -> beer_data

# Where the recipe is located. This assumes a Beersmith2 plaintext recipe. If
# you don't use Beersmith2 or don't want to include the recipe use the 
# Beer Stats Flyer -- No Beersmith.Rmd script.
# Pull in the recipe from Beersmith2
recipe <- "~/Google Drive/Beer Things/Recipes/Munich Helles.txt"
```

```{r include = FALSE}
# dplyr for processing data
if(require("dplyr")){
  print("dplyr is loaded correctly")
} else {
  print("trying to install dplyr")
  install.packages("dplyr")
  if(require("dplyr")){
    print("dplyr installed and loaded")
  } else {
    stop("could not install dplyr")
  }
}

# ggplot2 for graphing and scales for easy scaling
if(require("ggplot2")){
  print("ggplot2 is loaded correctly")
} else {
  print("trying to install ggplot2")
  install.packages("ggplot2")
  if(require("ggplot2")){
    print("ggplot2 installed and loaded")
  } else {
    stop("could not install ggplot2")
  }
}
if(require("scales")){
  print("scales is loaded correctly")
} else {
  print("trying to install scales")
  install.packages("scales")
  if(require("scales")){
    print("scales installed and loaded")
  } else {
    stop("could not install scales")
  }
}

# Lubridate for handling date formats easily
if(require("lubridate")){
  print("lubridate is loaded correctly")
} else {
  print("trying to install lubridate")
  install.packages("lubridate")
  if(require("lubridate")){
    print("lubridate installed and loaded")
  } else {
    stop("could not install lubridate")
  }
}

# readr for pulling in recipe file
require("readr")
if(require("readr")){
  print("readr is loaded correctly")
} else {
  print("trying to install readr")
  install.packages("readr")
  if(require("readr")){
    print("readr installed and loaded")
  } else {
    stop("could not install readr")
  }
}

# Knitr
require("knitr")
if(require("knitr")){
  print("knitr is loaded correctly")
} else {
  print("trying to install knitr")
  install.packages("knitr")
  if(require("knitr")){
    print("knitr installed and loaded")
  } else {
    stop("could not install knitr")
  }
}
```


```{r include = FALSE}
# Set names
names(beer_data) <- c("Timepoint", "SG", "Temp", "Color", "Beer", "Comment")

# Convert the Timepoint into an R readable time format
# Replace PST with the time zone the data was recorded in.
beer_data$Timepoint <- ymd_hms(beer_data$Timepoint, tz = "PST")

# Sort by date
beer_data <- arrange(beer_data, Timepoint)

# Convert Color and Beer name to Factor
beer_data$Color <- factor(beer_data$Color)
beer_data$Beer <- factor(beer_data$Beer)

# Scrape the beer name from the data for setting as the title
beer_name <- paste(as.character(beer_data$Beer)[nrow(beer_data)])

# Pull out the date range of fermentation for the title
date_range <- paste(format(min(beer_data$Timepoint), "%b %d, %Y"), "-",
      format(max(beer_data$Timepoint), "%b %d, %Y"))

# Style
style <- gsub(read_lines(recipe)[grepl(read_lines(recipe), pattern = "Style:")], pattern = "Style: ", replacement = "")
```

---
title: "`r beer_name` (`r style`)"
date: "`r date_range`"
geometry: margin=2cm
---

```{r include = FALSE}
# Drop any Gravity readings that are jumps off the local mean. This is to get
# rid of those readings where the Tilt jumps due to a bubble or being stirred.
# First we drop incomplete cases of SG since they won't be helpful to us.
beer_data <- beer_data[complete.cases(beer_data$SG), ]

# Then we fit a loess model to the data for a comparison.
loessMod <- loess(SG ~ as.numeric(Timepoint), data = beer_data, span = 0.20)
beer_data$SG_smooth <- as.vector(loessMod$fitted)

# Finally, we replace extreme values with the smoothed mean to normalize the data.
beer_data$SG <- ifelse(beer_data$SG >= beer_data$SG_smooth + .005, beer_data$SG_smooth, beer_data$SG)
beer_data$SG <- ifelse(beer_data$SG <= beer_data$SG_smooth - .005, beer_data$SG_smooth, beer_data$SG)

# Beer Statistics ---------------------------------------------------------
# Calculate the Original Gravity (OG) of the bear from a mean of the last
# 12 measurement points (first 3 hours). Can be extended for as much as desired.
select(beer_data, "SG") %>%
  slice(1:12) %>%
  colMeans() -> OG

# Convert OG into degrees Plato
Original_Plato = (-1 * 616.868) + (1111.14 * OG) - (630.272 * OG ^ 2) + (135.997 * OG ^ 3)

# Repeat for FG except take the first 12 rows.
select(beer_data, "SG") %>%
  slice((nrow(beer_data) - 11):nrow(beer_data)) %>%
  colMeans() -> FG

Final_Plato = (-1 * 616.868) + (1111.14 * FG) - (630.272 * FG ^ 2) + (135.997 * FG ^ 3)

# Calculate ABV and ABW. We use both the approximation formula which is
# fine for low alcohol beers. But, as alcohol content rises the accuracy of
# the fomula degrades. Both are presented for fun.
ABV_approximate <- (OG - FG) * 131.25
ABW <- 76.08 * (OG - FG) / (1.775 - OG)
ABV_specific <- ABW * (FG / 0.794)

# Calculate the apparent degree of fermentation (apparent attentuation)
ADF <- ((OG - FG) / (OG - 1)) * 100

# Calculate the real extract
Real_Extract = 0.188 * Original_Plato + 0.8192 * Final_Plato

# And the real attentuation
RA <- ((Original_Plato - Real_Extract) / Original_Plato) * 100

# Caloric content in 12oz and 22oz bottles
Calories_12oz <- (6.9 * ABW + 4.0 * (Real_Extract - 0.1)) * 3.54882 * FG 
Calories_22oz <- (6.9 * ABW + 4.0 * (Real_Extract - 0.1)) * 6.50618 * FG 

Color <- gsub(read_lines(recipe)[grepl(read_lines(recipe), pattern = "Estimated Color:")], pattern = "Estimated Color: ", replacement = "")

IBU <- gsub(read_lines(recipe)[grepl(read_lines(recipe), pattern = "Estimated IBU:")], pattern = "Estimated IBU: ", replacement = "")

# Format ingredients so they show up well
ingredients <- read_fwf(
  recipe,
  fwf_cols(
  Amount = 22,
  Ingredient = 41,
  Type = 14,
  Num = 9,
  Total = NA
  ),
  skip = (which(
  grepl(read_lines(recipe), pattern = "Ingredients")
  ) + 2),
  n_max = ((which(
  grepl(read_lines(recipe), pattern = "Mash Schedule")
  ) - 3) -
  (which(
  grepl(read_lines(recipe), pattern = "Ingredients")
  ) + 2))
  )

data.frame(select(ingredients, -Num)) -> display_recipe
```
\vspace{.5cm}
\center
ABV = `r round(ABV_approximate, 2)`\% to `r round(ABV_specific, 2)`\%; ABW = `r round(ABW, 2)`\%

IBUs = `r IBU`

Color = `r Color`

Original Gravity = `r paste(round(OG, 3))`; Original Plato = `r round(Original_Plato, 3)`

Final Gravity = `r round(FG, 3)`; Final Plato = `r round(Final_Plato, 3)`

Apparent Attenuation = `r round(ADF, 2)`%; Real Attentuation = `r round(RA, 2)`\%

Calories in 22oz Bottle = `r round(Calories_22oz, 0)`

\vspace{.5cm}

\Large Recipe

\normalsize
```{r, fig.align = 'left', echo = FALSE, comment = NULL, warning = FALSE, error = FALSE, message = FALSE, }
kable(display_recipe, align = "l")
```


<!-- ``` {r, echo = FALSE, comment = NULL, warning = FALSE, error = FALSE, message = FALSE, fig.width = 5, fig.height = 3, fig.align = "center"} -->
<!-- # Gravity by Time -->
<!-- ggplot(beer_data, aes(x = Timepoint, y = SG)) + -->
<!--   geom_line(aes(color = SG)) + -->
<!--   geom_smooth(color = "steelblue", se = FALSE) + -->
<!--   scale_color_gradient(low = "darkgreen", high = "lightgreen", guide = "none") + -->
<!--   scale_x_datetime(breaks = date_breaks("2 days"), labels = date_format("%b, %d")) + -->
<!--   scale_y_continuous(breaks = seq(from = min(beer_data$SG), to = max(beer_data$SG), by = .004)) + -->
<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1)) + -->
<!--   xlab("Time") + -->
<!--   ylab("Specific Gravity") + -->
<!--   ggtitle("Specific Gravity by Time Linegraph") -->
<!-- ``` -->

<!-- ``` {r echo = FALSE, comment = NULL, warning = FALSE, error = FALSE, message = FALSE, fig.width = 5, fig.height = 3} -->
<!-- # Temperature by Time -->
<!-- ggplot(beer_data, aes(x = Timepoint, y = Temp)) +  -->
<!--   geom_line(aes(color = Temp)) +  -->
<!--   geom_smooth(color = "steelblue", se = FALSE) + -->
<!--   scale_color_gradient(low = "blue", high = "red", guide = "none") +  -->
<!--   scale_x_datetime(breaks = date_breaks("2 days"), labels = date_format("%b, %d")) + -->
<!--   scale_y_continuous(breaks = seq(from = round(min(beer_data$Temp), 0), to = round(max(beer_data$Temp), 0), by = 3)) + -->
<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1)) + -->
<!--   xlab("Time") +  -->
<!--   ylab("Temperature (F)") +  -->
<!--   ggtitle("Temperature by Time Linegraph") -->
<!-- ``` -->